### 2014-11-18 docs/ms.Rmd
Integrated Kate's edits into ms.Rmd

@@

## Introduction
Since Thayer [-@thayer_concealing_1909], Poulton [-@poulton_colours_1890] and Wallace [-@wallace_natural_1891] generations of biologists have endeavored to explain the mechanisms and functions of animal and plant colouration. For scientists of this discipline, identifying best practice in measuring colour and colour pattern and in determining how other species see colour has been a great challenge. Over the past 150 years biologists, both independently and in collaboration with physicists and chemists, have developed a range of techniques used to measure and analyse biological colouration including spectrometry (also called spectroradiometry, spectrophotometry etc.), photography, and visual modeling.

The application of spectrometry to questions in ecology and evolution revolutionized the field [@dyck_determination_1966]. Before the widespread use of spectrometry, researchers assigned categorical colour descriptions, arbitrary rankings or scores [e.g. @aldrich_ecogeographic_1991; @bailey_latitudinal_1978; @wheelwright_colors_1985] or matched paint swatches or munsell chip ratings based on their own visual assessment [e.g. @burtt_analysis_1986; @key_temperature_1954]. Although no doubt applied carefully, these techniques limited progress in the field. For example, many researchers designed colour assessment methods specifically for their projects [e.g. @aldrich_ecogeographic_1991; @bailey_latitudinal_1978] rendering data by different researchers incomparable. This is especially problematic when different measurement techniques are used on the same species [e.g. the house finch _Carpodacus mexicanus_; @aldrich_origin_1978; @hill_female_1990]. Also, human vision, insensitive to ultraviolet wavelengths, is inherently different to most other animals such as avian and reptilian tetrachromats, and insect trichromats [@bennett_sexual_1994; @douglas_spectral_2014]. Human perception of colour is not only biologically irrelevant to most systems, but can lead to erroneous conclusions [@bennett_sexual_1994; @endler_measurement_1990].

Measuring reflectance spectra is now a reasonably inexpensive and convenient technique. Spectrometry allows the capture of quantitative and arguably repeatable data and has been rapidly taken up by researchers the field. However one drawback to this methodology, is the ease with which erroneous data can be collected if attention is not paid to finer details of spectrum capture. The angle from which spectra are taken [@santos_strong_2007], consistent calibration of light and dark standards, the distance between light source and object, the expectations of the researcher and the size, shape, glossiness or iridescence of the specimen [@andersson_quantifying_2006; @kemp_ornamental_2008; @meadows_quantifying_2011; @schaefer_birds_2008], can all contribute to potentially skewed data or alter results. Experience, guidance and objectivity are necessary to ensure data are consistent and representative of the subject. To ensure repeatability, the conditions under which data were collected must be reported exactly.

In conjunction with spectrometry, digital photography is increasingly being implemented in colour quantification. High-resolution cameras are now inexpensive and portable, allow instantaneous sampling of multiple colour patches, and can produce highly repeatable images [@mckay_use_2013]. The main challenge for gathering and using colour data from photographs is ensuring correct standardization methods, such as linearization, and consistent lighting [@stevens_using_2007]. Crucially, most digital cameras used in colour science are consumer products designed for human visual colour space, and hardware modifications for ultraviolet and infrared imaging are generally necessary [e.g. @stevens_colour_2013], making photography complementary, rather than superior to spectrometry.

@@

Sensory models typically attempt to describe the subjective perception of colour or brightness information as a function of an object’s reflectance, the ambient illumination, and a receiver’s sensory system [e.g. @endler_comparing_2005; @vorobyev_receptor_1998]. These models enable researchers to move beyond purely quantitative comparisons of reflectance spectra and adopt biologically relevant perspectives of signal receivers to test/formulate hypotheses. Although relatively easy to implement and potentially powerful, sensory models are built on multiple assumptions about the way in which stimuli are visually and cognitively processed that can dramatically shape the results of a given analysis [@lind_avian_2009; @pike_preserving_2012]. For example, Vorobyev et al.’s [-@vorobyev_receptor_1998; @vorobyev_colour_2001] widely used receptor-noise model requires researchers to specify the form of receptor quantum catches (i.e. raw versus log-transformed), the use of any colour constancy mechanism (e.g. von Kries), as well as the photoreceptor signal-to-noise ratios, their relative densities, and the type of noise predominating (i.e. neural and/or receptor noise). These data are only available for a small number of model species, such as honeybees, and so assumptions and extrapolations are necessary when models are applied to related species. Unless the biological rationale underlying every parameter choice is clearly understood, justified, and reported, such analyses may become ineffective, especially when comparing results between studies [@pike_preserving_2012].

Since the repeatability and transparency of science is paramount, it is imperative that we regularly and constructively reflect on our own practices [@alsheikh_public_2011; @sandve_ten_2013]. The potential for subjectivity in colour measurement and the mathematical opacity of visual models for the average colour scientist means that, as in any field, is it critical that methods are reported rigorously. Methods reporting should ideally have well defined standards such that sufficient detail is provided to allow comprehensive peer review, precise replication, cross-study comparison and data sharing [@alsheikh_public_2011]. The aims of this paper were to (a) review methods reporting in papers published on colour in 2013 from 22 top tier journals and (b) provide a set of recommendations for best-practice in reporting on the most common methods. Our hope is that this paper will provide a useful set of guidelines to improve the reproducibility and transparency of studies on colour.

>>>>

Article type: Commentary

Word count:

Number of figures:

Number of tables: 2

Data archiving:

\newpage

@@

\newpage

## Introduction
Generations of biologists have endeavored to explain the mechanisms and functions of animal and plant colouration [@thayer_concealing_1909; @poulton_colours_1890; @wallace_natural_1891]. For scientists of this discipline, identifying best practice in measuring colour and colour pattern has been a great challenge. For these biologists, identifying best practice in measuring color and color pattern has been a great challenge. Over the past 150 years, biologists have developed and adapted techniques that are now widespread, including spectrometry, photography, and visual modelling.

The application of spectrometry to biological questions revolutionized the field [@dyck_determination_1966]. Before its widespread use, researchers assigned categorical colour descriptions, arbitrary rankings or scores [e.g. @aldrich_ecogeographic_1991; @bailey_latitudinal_1978; @wheelwright_colors_1985] or matched paint swatches or munsell chip ratings based on their own visual assessment [e.g. @burtt_analysis_1986; @key_temperature_1954]. Although no doubt applied carefully, these techniques limited progress. For example, many researchers designed colour assessment methods specifically for their projects [e.g. @aldrich_ecogeographic_1991; @bailey_latitudinal_1978] rendering data by incomparable across studies. This is especially problematic when different measurement techniques are used on the same species [e.g. the house finch _Carpodacus mexicanus_; @aldrich_origin_1978; @hill_female_1990]. Likewise, because human vision is inherently different to most other animals [@bennett_sexual_1994; @douglas_spectral_2014], its use can lead to erroneous conclusions [@bennett_sexual_1994; @endler_measurement_1990].

Measuring reflectance spectra via spectrometry is now a reasonably inexpensive and convenient technique. Spectrometry allows the capture of quantitative and arguably repeatable data and has been rapidly taken up by researchers the field. However one drawback to this methodology, is the ease with which erroneous data can be collected if attention is not paid to finer details of spectrum capture. The angle from which spectra are taken [@santos_strong_2007], consistent calibration of light and dark standards, the distance between light source and surface, and the size, shape, glossiness or iridescence of the specimen can all potentially bias results [@andersson_quantifying_2006; @kemp_ornamental_2008; @meadows_quantifying_2011; @schaefer_birds_2008]. Experience, guidance and objectivity (e.g. double-blind experimental design) are necessary to ensure data are consistent and representative of the subject. Crucially, the conditions under which data were collected must be reported exactly to ensure repeatability.

Complementary to spectrometry, digital photography is increasingly being used to quantify biological colouration. High-resolution cameras are inexpensive,  and allow instantaneous, repeatable sampling of multiple colour patches [@mckay_use_2013]. The main challenge for gathering colour data from photographs are in ensuring correct standardization - such as linearization - and consistent lighting [@stevens_using_2007]. Most digital cameras used for the measurement of biological colouration are consumer products designed for the human visual system. Thus hardware modifications for ultraviolet and infrared imaging are generally necessary [e.g. @stevens_colour_2013].

@@

Sensory models typically attempt to describe the subjective perception of colour or brightness information as a function of an object’s reflectance, the ambient illumination, and a receiver’s sensory system [e.g. @endler_comparing_2005; @vorobyev_receptor_1998]. These models enable researchers to move beyond purely quantitative comparisons of reflectance spectra and adopt biologically relevant perspectives of signal receivers to formulate and test hypotheses.

Although relatively easy to implement and potentially powerful, sensory models are built on multiple assumptions about the way in which stimuli are visually and perceptually processed, and these can dramatically shape the results of a given analysis [@lind_avian_2009; @pike_preserving_2012]. For example, Vorobyev et al.’s [-@vorobyev_receptor_1998; @vorobyev_colour_2001] widely used receptor-noise model requires researchers to specify the form of receptor quantum catches (i.e. raw versus log-transformed), the use of any colour constancy mechanism (e.g. von Kries), as well as the photoreceptor signal-to-noise ratios, their relative densities, and the type of noise predominating (i.e. neural and/or receptor noise). These data are only available for a small number of species, such as humans and honeybees, and so assumptions and extrapolations are necessary when models are applied to other species. Unless the biological rationale underlying every parameter choice is clearly understood, justified, and reported, such analyses may become ineffective, especially when comparing results between studies [@pike_preserving_2012].

Since the repeatability and transparency of science is paramount, it is imperative that we regularly and constructively reflect on our own practices [@alsheikh_public_2011; @sandve_ten_2013]. The potential for subjectivity in colour measurement and the mathematical opacity of visual models for the average biologist means that, as in any field, is it critical that methods are reported rigorously. Methods reporting should ideally have well defined standards such that sufficient detail is provided to allow comprehensive peer review, precise replication, cross-study comparison and data sharing [@alsheikh_public_2011]. The aims of this paper were to (a) review methods reporting in papers published on colour in 2013 from 22 top tier journals and (b) provide a set of recommendations for best-practice in reporting on the most common methods. Our hope is that this paper will provide a useful set of guidelines to improve the reproducibility and transparency of studies on colour.
